---
title: "ABC output report."
output_document: html_document
params:
  datafile: "data/mixGauss3D.csv"
  input_variables: !r c("theta1", "theta2", "theta3") 
  prior_bounds: !r list(theta1=c(-10,10), theta2=c(-10,10), theta3=c(-10,10))
  bins: !r list(theta1=20, theta2=20, theta3=20)
  bandwidth: !r list(theta1=0.01, theta2=0.01, theta3=0.01)
---

```{r}
library(ks)
library(igraph)
library(mvtnorm)
library(tidyverse)
library(patchwork)
```

```{r}
datafile <- params$datafile
input_variables <- params$input_variables
prior <- as_tibble(params$prior_bounds)
dimension <- length(input_variables)
bins <- as_tibble(params$bins)
h <- as_tibble(params$bandwidth)
```

# Data

```{r}
post <- read_csv(datafile) %>% 
  select_at(c(input_variables, "weight")) %>%
  filter(weight > 0) %>%
  mutate(weight = weight / sum(weight))

thetas <- post %>% select_at(input_variables)
weights <- post %>% select(weight)
```

# Kernel density estimate

```{r}
source("kde.R")
```

```{r}
fhat <- kde_bounded_2(thetas, weights, bounds=prior, bins = bins,h = h)
```

Sanity check of the density estimates:

```{r}
fhat$result %>% summarize(sum(density*vcell)) 
```

# Marginales

```{r}
fhat_marginal <- function(thetas, weights, prior, bins, h) {
  bind_rows(
    map(
      names(thetas), 
      function(v) {
        f <- kde_bounded_2(select(thetas, v), weights, 
                            bounds = select(prior, v), 
                            bins = select(bins, v), 
                            h=select(h, v))
        f$result %>% select(v1=v, density) %>% mutate(v1_name=!!v)
      }
    )
  )
}

ggplot(fhat_marginal(thetas,weights,prior,bins,h)) +
  geom_line(aes(x=v1, y=density)) +
  facet_wrap("v1_name")
```

```{r}
fhat_marginal_2 <- function(thetas, weights, prior, bins, h) {
  variable_pairs <- t(combn(names(thetas), 2))
  marginals <- tibble(v1=variable_pairs[,1], v2=variable_pairs[,2])

  bind_rows(
    pmap(marginals, function(v1, v2) {
    f <- kde_bounded_2(thetas[c(v1, v2)], weights, bins=select(bins, v1, v2), 
      bounds=select(prior, v1, v2), h = select(h, v1, v2))
    tibble(v1 = f$result[[v1]], v2 = f$result[[v2]], v1_name=!!v1, v2_name=!!v2, density=f$result[["density"]])
    })
  )
}
```

```{r}
ggplot(fhat_marginal_2(thetas, weights, prior, bins, h)) + 
  geom_raster(aes(x=v1, y=v2, fill=density)) +
  # geom_contour(aes(x=v1, y=v2, z=density), colour="white") +
  facet_grid(rows=vars(v1_name), cols=vars(v2_name))
```


# Maximum density estimate of the model parameters


```{r}
most_likely <- function(fhat) {
  maxLikelyInd <- which.max(fhat$result$density)
  fhat$result[maxLikelyInd,]
}
```

```{r}
ml <- most_likely(fhat)
ml
```



# Is the posterior distribution flat or peaked?

Volume of the subspace GTau where the local probability is > tau, as a function of tau.

```{r}
vgtau <- function(fhat, tau) {
  fhat$result %>% filter(density >= tau) %>% summarize(sum(vcell)) %>% deframe()
}

vgtaus <- function(fhat, taus) {
  sapply(taus, function(x) { vgtau(fhat,x) })
}
```

```{r}
taus <- seq(0, max(fhat$result$density), length.out=50)
ggplot(data.frame(tau=taus, posterior=vgtaus(fhat, taus))) +
  geom_line(aes(x=tau, y=posterior)) +
  scale_x_reverse() +
  ylab("Volume")
```



# How many peaks are there and where are they?

Number of disjoint components of the subspace where the density is > tau, as a function of tau.

```{r}
cgtau <- function(fhat, tau) {
  if(tau <= 0) {
    c <- rep(1, nrow(fhat$result))
  } else {
    nodes <- (1:nrow(fhat$result))[fhat$result$density >= tau]
    g <- induced_subgraph(make_lattice(sapply(fhat$grid_borders, length)), nodes)
    c <- components(g)$membership
  }
  fhat$result %>% 
    filter(density >=tau) %>% 
    group_by(cluster = c) %>%
    filter(density == max(density)) %>%
    ungroup() %>%
    select_at(c(input_variables, "density")) %>%
    mutate(tau=tau)
}

taus <- seq(0, max(fhat$result$density), length.out=100)
cgtaus <- bind_rows(lapply(taus, function(x) cgtau(fhat, x)))

ggplot(cgtaus %>% group_by(tau) %>% summarize(clusters = n())) +
  geom_line(aes(x=tau, y=clusters)) +
  scale_x_reverse()
```

Highest position of each peak.

```{r}
cgtaus %>% group_by_at(input_variables) %>% summarize(density=max(density)) %>% ungroup() %>% arrange(desc(density)) 
```


# How close to the highest peak is the distribution?

Posterior probability of the parameters being within a sphere centered on the highest density point, as a function of its radius. 

```{r}
selected_points <- fhat$result %>% filter(density > 0)

# Distance of each point from the highest density
sp <- selected_points %>% select(input_variables)
repml <- matrix(rep(as.numeric(ml[,input_variables]), each=nrow(sp)), ncol=dimension)
dhd <- sqrt(rowSums((sp - repml) ** 2))

# Cumulative probability as a function of the radius
ppoint <- selected_points$density * selected_points$vcell
sortind <- order(dhd)
prad<- cumsum(ppoint[sortind])

ggplot(data.frame(radius = dhd[sortind], posterior = prad)) +
  geom_line(aes(x=radius, y=posterior))
```





# Model checking

## Posterior predictive checks: is the model a resonable explanation of the data?

```{r}
output_variables <- c("o1", "o2", "o3")

predThetas <- read_csv("data/posteriorPrediction.csv") %>% 
  select_at(input_variables)
predOut <- read_csv("data/posteriorPrediction.csv") %>% 
  select_at(output_variables)
```

### Marginals

```{r}
bounds_out <- tibble(o1=c(-10,10), o2=c(-10,10), o3=c(-10,10))
bins_out <- tibble(o1=20, o2=20, o3=20)
h_out <- tibble(o1=0.01, o2=0.01, o3=0.01)
fhat_pred <- kde_bounded_2(predOut, 1, bounds=bounds_out, bins = bins_out,h = h_out)

fhat_pred$result %>% summarize(sum(density * vcell))
```

```{r}
observed <- bind_rows(tibble(o1 = 0, o2 = 0, o3 = 0))

ggplot(fhat_marginal(predOut,1,bounds_out,bins_out,h_out)) +
  geom_line(aes(x=v1, y=density)) +
  facet_wrap("v1_name") + 
  geom_vline(aes(xintercept=v1), gather(observed, "v1_name", "v1"), colour="red") +
  ylim(0,NA)
```

```{r}
ggplot(fhat_marginal_2(predOut, 1, bounds_out, bins_out, h_out)) + 
  geom_raster(aes(x=v1, y=v2, fill=density)) +
  # geom_contour(aes(x=v1, y=v2, z=density), colour="white") +
  facet_grid(rows=vars(v1_name), cols=vars(v2_name)) +
  geom_point(aes(x=v1, y=v2), 
    as_tibble(t(combn(names(predOut), m=2))) %>% rename(v1_name=1,v2_name=2) %>% mutate(v1=as_vector(observed[v1_name]), v2=as_vector(observed[v1_name])), colour="red", size=3)
```


### Credible set.

Probability of the subspace GTau where the local probability is > tau, as a function of tau.

```{r}
pgtau <- function(fhat, tau) {
  fhat$result %>% filter(density > tau) %>% summarize(sum(density * vcell)) %>% deframe()
}

pgtaus <- function(fhat, taus) {
  sapply(taus, function(x) { pgtau(fhat,x) })
}

```

Value of tau such that the set PGtau has probability 0.95.

```{r}
max_den_pred <- max(fhat_pred$result$density)
taus_pred <- seq(0, max_den_pred, length.out=50)
pgtaus_pred <- tibble(tau=taus_pred, posterior=pgtaus(fhat_pred, taus_pred)) %>% 
  bind_rows(tibble(tau=Inf, posterior=0))

tau_pg95 <- pgtaus_pred %>% arrange(posterior) %>%
  slice(c(last(which(posterior <= 0.95)), first(which(posterior > 0.95))))
tau_pg95 <- approx(tau_pg95$posterior, tau_pg95$tau, 0.95)$y
 
tau_pg95
```

Estimated density at the observed point.

```{r}
fhat_observed <- kde_bounded_2(predOut, 1, bounds=bounds_out, h = h_out, evalp=observed)$result$density
fhat_observed
```

```{r}
pg_fhat_observed <- pgtaus_pred %>% arrange(tau) %>%
  slice(last(which(tau <= fhat_observed))) %>% 
  select(posterior) %>% deframe()

ggplot(pgtaus_pred) +
  geom_line(aes(x=tau, y=posterior)) +
  scale_x_reverse() +
  annotate("segment", x = Inf, y = pg_fhat_observed, xend = fhat_observed, yend=pg_fhat_observed, colour = "red") +
  annotate("rect", xmin = tau_pg95, xmax = Inf, ymin = -Inf, ymax = 0.95, alpha = .2) + 
  annotate("segment", x = fhat_observed, y = pg_fhat_observed, xend = fhat_observed, yend=-Inf, colour = "red") +
  ylab("Probability")
```



