---
title: "ABC output report."
output_document: html_document
params:
  datafile: "mixGauss.csv"
  input_variables: !r c("theta0", "theta1") 
  prior_bounds: !r list(min=c(-10,-10), max=c(10,10))
  bandwidth: !r c(0.005, 0.005)
---

```{r eval=FALSE}
params <- list(datafile = "mixGauss.csv",
              input_variables = c("theta0", "theta1"),
              prior_bounds = list(min=c(-10,-10), max=c(10,10)),
              bandwidth = c(0.001, 0.001))
```

```{r}
library(ks)
library(igraph)
library(mvtnorm)
library(tidyverse)
library(patchwork)
```

```{r}
datafile <- params$datafile
input_variables <- params$input_variables
prior <- as_tibble(params$prior_bounds)
dimension <- length(input_variables)
h <- params$bandwidth
```

# Data

```{r}
post <- read_csv(params$datafile) %>% 
  select_at(c(params$input_variables, "weight")) %>%
  filter(weight > 0) %>%
  mutate(weight = weight / sum(weight))

thetas <- post %>% select_at(params$input_variables)
weights <- post %>% select(weight)
```

# Kernel density estimate

```{r}
source("kde.R")
```

```{r}
#fhat <- fhat_bounded(thetas_coop, weights_coop, bounds=data.frame(min=c(0,0,0), max=c(1,1,1)), bins = rep(50,ncol(thetas_coop)))
#fhat <- fhat(thetas_coop, weights_coop, bins = rep(50,ncol(thetas_coop)))
# fhat_mg <- fhat(thetas_mg, weights_mg, bins = rep(50,ncol(thetas_mg)))
fhat <- kde_bounded_2(thetas, weights, bounds=prior, bins = rep(30,ncol(thetas)),h = h)
```

Sanity check of the density estimates:

```{r}
fhat$result %>% summarize(sum(density*vcell)) %>% deframe()
```

# Marginales

```{r}
fhat_marginal <- map2(input_variables, 1:dimension, function(v,i) 
    kde_bounded_2(thetas[v], weights, bins=100, bounds=slice(prior, i), h=h[[i]])) %>% 
  set_names(input_variables)

marginal_plots <- imap(fhat_marginal, function(marginal,variable_name)
  ggplot(marginal$result) + geom_line(mapping=aes_string(x=variable_name, y="density")))


marginal_dotplots <- imap(input_variables, function(variable_name, variable_index)
  ggplot(thetas) + geom_dotplot(aes_string(x=variable_name), binwidth=(max(thetas[[variable_name]]) - min(thetas[[variable_name]])) / 300))

reduce(c(marginal_plots,marginal_dotplots), `+`) + 
  if(length(marginal_plots) > 1) {plot_layout(ncol=min(c(dimension,4)))}
```

```{r}
variable_pairs <-t(combn(input_variables, 2))
indices_pairs <- t(combn(1:dimension, 2))
marginals <- tibble(v1=variable_pairs[,1], v2=variable_pairs[,2], i1=indices_pairs[,1], i2=indices_pairs[,2])
fhat_marginal_2 <- pmap(marginals, function(v1, v2, i1, i2) 
  kde_bounded_2(thetas[c(v1, v2)], weights, bins=50, bounds=slice(prior, i1, i2), h = h[c(i1, i2)]))
```

```{r}
plot_marginal_2 <- function(fhat_marg, v1, v2) {
  ggplot(fhat_marg$result) + 
    geom_raster(aes_string(x=v1, y=v2, fill="density")) +
    geom_contour(aes_string(x=v1, y=v2, z="density"), colour="white") 
}

marginal_2_plots <- pmap(list(fhat_marg=fhat_marginal_2, v1=variable_pairs[,1], v2=variable_pairs[,2]), plot_marginal_2)

reduce(marginal_2_plots, `+`) + if (length(marginal_2_plots) > 1) {plot_layout(ncol=2)}
```



# Maximum density estimate of the model parameters


```{r}
most_likely <- function(fhat) {
  maxLikelyInd <- which.max(fhat$result$density)
  fhat$result[maxLikelyInd,]
}
```

```{r}
ml <- most_likely(fhat)
ml
```



# Is the posterior distribution flat or peaked?

Volume of the subspace GTau where the local probability is > tau, as a function of tau.

```{r}
vgtau <- function(fhat, tau) {
  #sum(fhat$result$density >= tau * fhat$vcell)
  fhat$result %>% filter(density >= tau) %>% summarize(sum(density * vcell)) %>% deframe()
}

vgtaus <- function(fhat, taus) {
  sapply(taus, function(x) { vgtau(fhat,x) })
}
```

```{r}
taus <- seq(0, max(fhat$result$density), length.out=50)
ggplot(data.frame(tau=taus, posterior=vgtaus(fhat, taus))) +
  geom_line(aes(x=tau, y=posterior)) +
  scale_x_reverse() +
  ylab("Volume")
```


# How many peaks are there and where are they?

Number of disjoint components of the subspace where the density is > tau, as a function of tau.

```{r}
cgtau <- function(fhat, tau) {
  if(tau <= 0) {
    c <- rep(1, nrow(fhat$result))
  } else {
    nodes <- (1:nrow(fhat$result))[fhat$result$density >= tau]
    g <- induced_subgraph(make_lattice(sapply(fhat$grid_borders, length)), nodes)
    c <- components(g)$membership
  }
  fhat$result %>% 
    filter(density >=tau) %>% 
    group_by(cluster = c) %>%
    filter(density == max(density)) %>%
    ungroup() %>%
    select_at(c(input_variables, "density")) %>%
    mutate(tau=tau)
}

taus <- seq(0, max(fhat$result$density), length.out=100)
cgtaus <- bind_rows(lapply(taus, function(x) cgtau(fhat, x)))

ggplot(cgtaus %>% group_by(tau) %>% summarize(clusters = n())) +
  geom_line(aes(x=tau, y=clusters)) +
  scale_x_reverse()
```

Highest position of each peak.

```{r}
cgtaus %>% group_by_at(input_variables) %>% summarize(density=max(density)) %>% ungroup() %>% arrange(desc(density)) 
```


# How close to the highest peak is the distribution?

Posterior probability of the parameters being within a sphere centered on the highest density point, as a function of its radius. 

```{r}
selected_points <- fhat$result %>% filter(density > 0)

# Distance of each point from the highest density
sp <- selected_points %>% select(input_variables)
repml <- matrix(rep(as.numeric(ml[,input_variables]), each=nrow(sp)), ncol=dimension)
dhd <- sqrt(rowSums((sp - repml) ** 2))

# Cumulative probability as a function of the radius
ppoint <- selected_points$density 
sortind <- order(dhd)
prad<- cumsum(ppoint[sortind])

ggplot(data.frame(radius = dhd[sortind], posterior = prad)) +
  geom_line(aes(x=radius, y=posterior))
```





### Model checking ###

## 1. Is the model a resonable explanation of the data?




## 2. Equifinality
